---
title: "Olympics analysis"
author: "Tamas Nagy"
date: "10/20/2021"
output: 
  html_document:
   theme: spacelab
   code_download: true
   toc: true
   toc_float: true
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)

theme_set(theme_light())

```
# Summary

In this project, I will explore the olympics 

# Read data

We can read the data directly from Github.
```{r}
olympics <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

```

Create a new BMI variable
```{r}
olympics <-
    olympics %>% 
    mutate(bmi = weight / (height/100)^2)
```

# Some descriptive statistics about the Olympics. 

How many olympics have been between 1896-2016?
```{r}
olympics %>% 
    distinct(games, .keep_all = TRUE) %>% 
    count(season)
```

Olympic athletes
```{r}
olympics %>% 
    count(season)
```

Olympic cities
```{r}

olympics %>% 
    distinct(games, .keep_all = TRUE) %>% 
    count(city, sort = TRUE)

```

# Did the size of athletes change over time?

Sample 10,000 random athletes, and check how height and weight changed over the years by sex.

```{r}
# This is needed to keep the result the same even after random sampling
set.seed(123)

olympics %>% 
    drop_na(height, weight) %>% 
    sample_n(10000) %>% 
    pivot_longer(cols = height:weight,
                 names_to = "variable",
                 values_to = "value") %>% 
    ggplot() +
    aes(x = year, y = value, color = sex) +
    geom_jitter(size = 1, height = 0, alpha = .5) +
    geom_smooth(method = "lm") +
    facet_wrap(~variable, 
               nrow = 2,
               scales = "free_y") +
    labs(x = NULL,
         y = "Height (cm) or Weight (kg)",
         title = "Height and weight of olympic athletes over the years (10k random athletes)",
         subtitle = "Both males and females show an increase over time.")
    

```

# Average age of athletes who won a medal over time

```{r}
olympics %>% 
    drop_na(medal, age) %>% 
    group_by(year) %>% 
    summarise(avg_age = mean(age),
              se_age = sd(age)/sqrt(n())) %>% 
    ggplot() +
    aes(x = year,
        y = avg_age,
        ymin = avg_age - se_age,
        ymax = avg_age + se_age) +
    geom_ribbon(alpha = .2) +
    geom_line(size = 1.2, color = "red") +
    labs(title = "Average age of athletes who won a medal",
         subtitle = "Grey ribbon represents standard error.",
         y = "Age (years)",
         x = NULL)
    

```

# Which sports were included in the Olympics the most often?

```{r}
olympics %>% 
    distinct(year, season, sport) %>% 
    count(season, sport, sort = TRUE) %>% 
    # To sort by frequency, 
    mutate(sport = fct_reorder(sport, n, sum)) %>% 
    ggplot() +
    aes(y = sport, x = n, fill = sport) +
    geom_col(show.legend = FALSE) +
    scale_fill_viridis_d(option = "magma") +
    facet_wrap(~season, scales = "free_y") +
    labs(y = NULL,
         title = "Nr of times when sport was included",
         caption = "Figure Skating and Ice Hockey were also part of the Summer games in 1908 London, and 1920 Antwerpen", 
         x = "Nr")
    
# Which summer olympics included Figure Skating and Ice Hockey?
olympics %>% 
    filter(season == "Summer", sport %in% c("Figure Skating", "Ice Hockey")) %>% 
    distinct(games, city)

```

# Countries with the most athletes

```{r}
library(ggmap)
library(countrycode)
library(ggthemes)

olympics_world <-
    map_data(map = "world") %>% 
    mutate(noc = countryname(sourcevar = region, destination = "ioc")) %>% 
    left_join(count(olympics, noc), by = "noc") %>% 
    filter(region != "Antarctica")

olympics_world %>% 
    ggplot() +
    aes(map_id = region, long, lat, label = n) +
    geom_map(map = olympics_world,
             aes(fill =  log10(n))) +
    theme_map() +
    coord_quickmap() +
    labs(title = "The number of athletes sent to all olympics by each country.",
         subtitle = "There are a few countries that sent tens of thousands of athletes, but also a few that sent only tens at maximum.",
         x = NULL, y = NULL,
         fill = "Magnitude of athletes (10^x)")

```
